---
title: "General Discussion on Probit Priors for Likert Data Analysis"
author: "Anders Sundelin"
date: "2025-05-01"
output: html_document
params: 
    cache: "../.cache"
    output: "../output"
    reloo: FALSE
    cores: 2
    threads: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(forcats)
library(stringr)
library(lubridate)
library(bookdown)
library(ggformula)
library(brms)
library(bayesplot)
library(tidybayes)
library(ggpubr)
```

```{r ingest}
source("ingest_survey_data.R")
raw <- ingest_survey_data("../data/TrainingDayResponses.xlsx")
jaeger <- jaeger_data(raw) 
```
```{r}
jaeger_usable <- get_jaeger_usable(jaeger)
jaeger_ease_of_use <- get_jaeger_ease_of_use(jaeger)
jaeger_accessible <- get_jaeger_accessible(jaeger)
jaeger_intent <- get_jaeger_intend_use(jaeger)
```

# Prior Selection 7-level Likert Analysis

Discussing what priors are suitable for analysing a 7-level Likert scale, where all options are possible.

Calculating thresholds for similar probability of each of the seven responses (given probit regression, i.e. $\Phi$ as cumulative distribution function).
```{r}
tibble(rating = 1:7) |>
  mutate(proportion = 1/7) |>
  mutate(cumulative_proportion = cumsum(proportion)) |>
  mutate(right_hand_threshold = qnorm(cumulative_proportion))
```

Unfortunately, `brms` do not yet support Dirichlet priors, so we make do with six $\mathcal{N}(\tau_i, 1)$ distributions.
Because of overlapping tails, this means that there is slightly more probability in the middle of the scale, but this is also quite typical of responses.

## Comparison with default priors

```{r}
d <- jaeger_usable |> select(usable)
formula <- usable | thres(6) ~ 1
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6))
validate_prior(prior = priors,
               data=d, 
               family=cumulative(probit),
               formula=formula)
```

```{r ppc}
(left_p <- pp_check(brm(
  data = d,
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,    
  chains = CHAINS,  
  cores = CORES,
  seed = 1
 ), type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
)
```

No divergent transitions using those priors, and we see that all Likert levels are quite likely.
How do these priors compare to the default `student_t(3, 0, 2.5)`?

```{r}
(right_p <- pp_check(brm(
  data = d,
  family = cumulative(probit),
  formula=formula,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  warmup = 1000, 
  iter  = ITERATIONS,    
  chains = CHAINS,
  cores = CORES,
  seed = 1
), type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
)
```

```{r}
ggarrange(left_p + ggtitle("Six N(tau,1) priors"), right_p + ggtitle("Default prior, Student_t(3, 0, 2.5)"), ncol=2, widths=c(1,1))
```


Compared to the default `brms` priors, our priors are more tolerant to values at the end of the scale (i.e., 1, 2, 6, 7).
This is visible as the expected value (dot) is clearly visible.
For our data, which has lot of values tending positive (towards 1), this is appropriate.
We would expect more data towards the end of the scale.

# Standard deviation priors.

Visualization of the exponential distribution, which we use for standard deviations.

```{r}
ggplot(data.frame(x=seq(0,4,by=0.1))) + theme_bw() +
  stat_function(aes(x), fun=dexp, col="#2b83ba", linewidth=1, args=(mean=1)) +
  stat_function(aes(x), fun=dexp, col="#fdae61", linewidth=1, args=(mean=2)) + 
  stat_function(aes(x), fun=dexp, col="#d7191c", linewidth=1, args=(mean=5)) + 
  stat_function(aes(x), fun=dexp, col="#abdda4", linewidth=1, args=(mean=10)) 
```

For a probit regression using $\mathcal{N}(0, 1)$, values outside [-3,3] is very unlikely (as >99% probability is contained with those six standard deviation).
This means that `exp(1)` is quite wide, and cause the sampler to go too much astray.
On the other hand, `exp(10)` could be seen as too narrow, so we settle for the middle, and use `exp(5)` for most of our standard deviation priors.

