---
title: "Survey Analysis GenAI"
author: "Anders Sundelin"
date: "2025-05-04"
output: html_document
params:
    cache: "../.cache"
    output: "../output"
    reloo: FALSE
    cores: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(forcats)
library(stringr)
library(lubridate)
library(bookdown)
library(ggformula)
library(brms)
library(bayesplot)
library(tidybayes)
```

# AI Stream
```{r validate-params}
stopifnot(dir.exists(params$cache), dir.exists(params$output))
```


```{r ingest}
source("ingest_survey_data.R")
raw <- ingest_survey_data("../data/TrainingDayResponses.xlsx")
ai <- ai_data(raw)
MODEL_PREFIX <- "genai"
```

## EDA, visualizations
```{r}
ai_usable <- get_ai_usable(ai)
ai_ease_of_use <- get_ai_ease_of_use(ai)
ai_accessible <- get_ai_accessible(ai)
ai_intent <- get_ai_intend_use(ai)
```

```{r}
rbind(ai_usable |> mutate(trait="usable", rate=usable) |> select(trait, site, rate),
      ai_ease_of_use |> mutate(trait="ease-of-use", rate=ease_of_use) |> select(trait, site, rate),
      ai_accessible |> mutate(trait="accessible", rate=accessible) |> select(trait, site, rate),
      ai_intent |> mutate(trait="intent-to-use", rate=intent) |> select(trait, site, rate)
      ) |> group_by(trait) |> ggplot(aes(x=as.numeric(rate), fill=site)) +
    geom_histogram(stat="count", position="dodge") +
  scale_x_continuous(name="rating", breaks = 1:7, labels = c("Extremely\nLikely", "Quite\nLikely", "Slightly\nLikely", "Neutral", "Slightly\nUnlikely", "Quite\nUnlikely", "Extremely\nUnlikely") ) +
  facet_wrap(~trait, ncol=2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Usability

## Usability, Population-level intercept only

```{r}
d <- ai_usable |> select(usable)
formula <- usable | thres(6) ~ 1
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)

```

The warning about the default prior not being used can safely be ignored.
We specigy our own priors for all parameters, and do not rely on `brms` default priors.

### Prior predictive checks

```{r m1_ppc}
M1_ppc <- brm(
  data = d,
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  seed = 1
)
```
No divergent transitions, all chains sampled OK

```{r}
pp_check(M1_ppc, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```
All outcomes are quite likely, with the middle ones having the highest likelihood (due to our use of separated normal distrubutions, the tails add up).

### Sampling
```{r m1}
M1 <- brm(
  data = d,
  file = modelfile("M1"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- M1
```

No divergent transitions. Let's do the posterior checks

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

The chains seem to mix well. The disc is flat as the width (variance) of the latent variable is fixed in this model.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Some $N_{eff}$ values are slightly below 0.5, but well above 0.1

```{r loo-m1}
( loo <- loo(M1) )
plot(loo)
```

No problems with the estimated LOO values, though one observation is above 0.6 (but still below the threshold 0.7).

### Posterior predictive checks

```{r}
pp_check(M1, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

Our model fits the data well.

```{r}
summary(M1)
```

```{r}
plot_M1_latent_distribution(M1, "usable")
```
```{r}
plot_M1_posterior_mean(M1, "usability", usable_mean)
```
```{r}
ai_usable |> summarise(mean(as.numeric(usable)))
```

Sample mean is visible in the figure as a dashed line (slightly closer to SL than QL, as the numeric value specifies)

## Usability, per site

```{r}
d <- ai_usable |> select(usable, site)
formula <- usable | thres(6) ~ 1 + site
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r m2_ppc}
M2_ppc <- brm(
  data = d,
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- M2_ppc
```

```{r}
pp_check(M2_ppc, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)

```

Priors looking good, all values are present.

```{r m2}
M2 <- brm(
  data = d,
  file = modelfile("M2"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- M2
```
No divergent transitions, all sampled correctly.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

All caterpillar plots look allright, chains mix correctly.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

One $N_{eff}$ value below 0.5, but most are fine.

```{r}
pp_check(m, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)

```

```{r}
plot_M2_latent_distribution(M2, "usability")

```
```{r}
plot_M2_posterior_mean(M2, "usability", usable_mean)
```

```{r}
d |> group_by(site, usable) |> tally()
```


# Multilevel, reguralized per respondent

```{r}
d <- ai_usable |> select(usable, site, id)
formula <- usable | thres(6) ~ 1 + site + (1 | id)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(exponential(5), class = sd))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)

```

```{r m3_ppc}
M3_ppc <- brm(
  data = d,
  family = cumulative(probit),
  formula = formula,
  prior = priors,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)

```

Two divergent transitions, but otherwise the priors look allright.

```{r}
pp_check(M3_ppc, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
m <- M3_ppc
```

These priors are looking good, all outcomes are possible
How does the groups look?

```{r}
pp_check(M3_ppc, type = "bars_grouped", group="id",
         ndraws = 500, size = 1/2, fatten = 3/2)

```

Grouped PPCs are also looking good, we should be able to trust these priors.

```{r m3}
M3 <- brm(
  data = d,
  file = modelfile("M3"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- M3
```

All chains sampled correctly.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

These caterpillars look good, and the `disc` continues to be flat, as it is fixed at 1.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

$\hat{R}$ are looking good, and although some $N_{eff}$ values are low, they range between 0.3 and 1.0.

```{r}
pp_check(m, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(m, type = "bars_grouped", group="id",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
plot_M2_latent_distribution(m, "usability")
```

Total overlap of Europe and Indian samples.

```{r}
plot_M2_posterior_mean(m, "usability", usable_mean)
```

Slightly wider distribution of expected value for Europe, even incorporating Neutral in the 95% probablerange.

# Multilevel site, pooling on respondent, variable variance (disc)

```{r}
d <- ai_usable |> select(usable, site, id)
formula <- bf(usable | thres(6) ~ 1 + site + (1 | id)) +
           lf(disc              ~ 0 + site, cmc = FALSE)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(normal(0, log(2)/2), class = b, dpar=disc),
            prior(exponential(5), class = sd)
            )
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r m4_ppc}
M4_ppc <- brm(
  data = d,
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  sample_prior = "only",
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
m <- M4_ppc
```

```{r}
pp_check(M4_ppc, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

These priors both sample well, and allows all possible values (1-7). Like the other models, the priors place most probability on the middle range of scores.

```{r m4}
M4 <- brm(
  data = d,
  file = modelfile("M4"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
m <- M4
```

This model is quite complex, so we have to adjust the step size (adapt_delta) and the initial value.
Normally, Stan draws initial values from a uniform distribution between [-2,2].
When using `cmdstanr`, we are a bit limited as regards to what `brms` allows (`init_r` does not seem to be supported)
So, here, we just set the initial value to the middle point of the default, i.e. `0`.
Sampling will continue from there, in various directions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

These caterpillars look good, and the `disc` continues to be flat, as it is fixed at 1.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

All $N_{eff}$ values are above 0.3, quite good.

```{r loo-m4}
loo <- loo(M4)
loo
plot(loo)
```

One bad LOO value, right on the limit 0.7.

Setting the RELOO option will run the exact Pareto calculation (this is not dune by default, to save run time).

```{r reloo-m4, eval=params$reloo}
reloofile <- paste0(modelfile("reloo-M4"), ".rds")
if (file.exists(reloofile)) {
    (reloo <- readRDS(reloofile))
} else {
    Sys.time()
    (reloo <- reloo(M4, loo, chains=CHAINS, cores=CORES) )
    Sys.time()
    saveRDS(reloo, reloofile)
}
```
Note that the reloo results are cached, and included in the docker image by default.
If you want to reproduce the reloo manually, set a separate cache dir (and mount it into the docker image), and enable the RELOO environment variable.

```{r plot_reloo-m4, eval=params$reloo}
plot(reloo)
```

Running reloo, which does exact LOO-CV (leaving out one problematic observation at a time) reveals that all Pareto-k estimates (the highest is less than 0.6) are below 0.7, so we should be able to use the model inferences.


```{r}
summary(M4)
```

```{r}
plot_M4_latent_distribution(M4, "usability")
```
```{r}
plot_M4_posterior_mean(M4, "usability-varying", usable_mean, limits=c(1,7))
```


# Ease-of-use


## Ease-of-use, intercept-only
```{r}
ai_ease_of_use <- get_ai_ease_of_use(ai)
```


```{r e1}
d <- ai_ease_of_use |> select(ease_of_use)
formula <- ease_of_use | thres(6) ~ 1
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
E1 <- brm(
  data = d,
  file = modelfile("E1"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- E1
```
No divergent transitions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

The chains seem to mix well, and the `disc` is flat.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good values, lowest $N_{eff}$ value is about 0.4.

```{r loo-e1}
loo <- loo(E1)
loo
plot(loo)
```

All Pareto-k values well below the limit for this simple model.

```{r}
pp_check(E1, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

No problems with the posterior checks or the LOO-CV estimate.

```{r}
plot_M1_latent_distribution(E1, "ease of use")
```

```{r}
pp_check(E1, type = "hist", ndraws = 8, binwidth = 1) +
  scale_x_continuous(breaks = 1:7) +
  ggtitle("Example draws from the posterior")
```

```{r}
plot_M1_posterior_mean(E1, "ease of use",  ease_of_use_mean)
```
```{r}
ai_ease_of_use |> summarise(mean(as.numeric(ease_of_use)))
```

## Ease-of-use, per site

```{r}
d <- ai_ease_of_use |> select(ease_of_use, site)
formula <- ease_of_use | thres(6) ~ 1 + site
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r e2}
E2 <- brm(
  data = d,
  file = modelfile("E2"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- E2
```

No divergent transitions

```{r}
pp_check(m, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```


```{r}
plot_M2_latent_distribution(E2, "ease of use")
```

Very similar distribution for the two sites.

```{r}
plot_M2_posterior_mean(E2, "ease of use",  ease_of_use_mean)
```

Slightly wider 95% credible interval for the European participants.

## Ease-of-use, per site, regularized per respondent

```{r}
d <- ai_ease_of_use |> select(ease_of_use, site, id)
formula <- ease_of_use | thres(6) ~ 1 + site + (1 | id)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(exponential(5), class = sd))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r e3}
E3 <- brm(
  data = d,
  file = modelfile("E3"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- E3
```

No divergent transitions

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

Caterpillar plots look OK

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is over 0.3.

```{r}
pp_check(m, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(m, type = "bars_grouped", group="id", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
plot_M2_latent_distribution(E3, "ease-of-use")
```

Similar distributions, slightly more separated that the non-grouped model E2.

```{r}
plot_M2_posterior_mean(E3, "ease-of-use",  ease_of_use_mean)
```

## Ease-of-use, pooled on respondent, varying variance per site
```{r}
d <- ai_ease_of_use |> select(ease_of_use, site, id)
formula <- bf(ease_of_use | thres(6) ~ 1 + site + (1 | id)) +
           lf(disc                   ~ 0 + site, cmc = FALSE)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(normal(0, log(2)/2), class = b, dpar=disc),
            prior(exponential(5), class = sd)
            )
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r e4}
E4 <- brm(
  data = d,
  file = modelfile("E4"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
m <- E4
```

No divergent transitions, all chains sampled well

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

These caterpillars look good, and the `b_disc_siteIndia` also mixes well.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is above 0.3

```{r loo-e4}
( loo <- loo(E4) )
plot(loo)
```

All Pareto-k values are well below the 0.7 limit, the highest around 0.4.

```{r}
summary(E4)
```

```{r}
plot_M4_latent_distribution(E4, "ease-of-use-varying")
```

Here we clearly see that the Indian respondents are tighter (less variance).
Which is weird, as their answers have more variance.
Perhaps the fewer responses in Europe give more impact to the priors?

```{r}
plot_M4_posterior_mean(E4, "ease-of-use-varying",  ease_of_use_mean, limits=c(1,7))
```

The predictions are very similar for the two sites, anyway

# Accessibility


## Accessibility, intercept-only
```{r}
ai_accessibility <- get_ai_accessible(ai)
```


```{r}
d <- ai_accessibility |> select(accessible)
formula <- accessible | thres(6) ~ 1
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r a1}
A1 <- brm(
  data = d,
  file = modelfile("A1"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- A1
```

No divergent transitions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

The chains seem to mix well, and the disc is flat.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is about 0.4.

```{r loo-a1}
( loo <- loo(A1) )
plot(loo)
```

One Pareto-k value around 0.6, but all are below 0.7.

```{r}
pp_check(A1, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

No problems with LOO-CV or the posterior checks

```{r}
plot_M1_latent_distribution(A1, "accessible")
```

```{r}
pp_check(A1, type = "hist", ndraws = 8, binwidth = 1) +
  scale_x_continuous(breaks = 1:7) +
  ggtitle("PP-check with the pp_check() function")
```

```{r}
plot_M1_posterior_mean(A1, "accessible", accessible_mean)
```

```{r}
ai_accessibility |> summarise(mean(as.numeric(accessible)))
```

Sample mean is closer to QL, as shown in the figure.

## Accessibility, per site

```{r}
d <- ai_accessibility |> select(accessible, site)
formula <- accessible | thres(6) ~ 1 + site
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r a2}
A2 <- brm(
  data = d,
  file = modelfile("A2"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- A2
```

Chains mix well

```{r}
pp_check(m, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)

```

Posterior predictions look representative

```{r}
plot_M2_latent_distribution(A2, "accessible")
```

Very similar distributions

```{r}
plot_M2_posterior_mean(A2, "accessible", accessible_mean)
```

## Accessibility, per site, regularized per respondent

```{r}
d <- ai_accessibility |> select(accessible, site, id)
formula <- accessible | thres(6) ~ 1 + site + (1 | id)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(exponential(5), class = sd))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r a3}
A3 <- brm(
  data = d,
  file = modelfile("A3"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- A3
```

No divergent transitions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

Chains mix well

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

OK diagnostics, lowest $N_{eff}$ is about 0.25.

```{r}
pp_check(m, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(m, type = "bars_grouped", group="id", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
plot_M2_latent_distribution(A3, "accessible")
```

Very similar distributions.

```{r}
plot_M2_posterior_mean(A3, "accessible", accessible_mean)
```

## Accessibility, pooled on respondent, varying variance per site

```{r}
d <- ai_accessible |> select(accessible, site, id)
formula <- bf(accessible | thres(6) ~ 1 + site + (1 | id)) +
           lf(disc                   ~ 0 + site, cmc = FALSE)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(normal(0, log(2)/2), class = b, dpar=disc),
            prior(exponential(5), class = sd)
            )
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r a4}
A4 <- brm(
  data = d,
  file = modelfile("A4"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
m <- A4
```

No divergent transitions, or starting point collapses, when we fix the step size and the init.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

Chains mix well.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is abit above 0.25 now.

```{r loo-a4}
( loo <- loo(A4) )
plot(loo)
```

Two observations with high Pareto-k values, one even above 1.
Setting the RELOO option will run the exact Pareto calculation (this is not dune by default, to save run time).

```{r reloo-a4, eval=params$reloo}
reloofile <- paste0(modelfile("reloo-A4"), ".rds")
if (file.exists(reloofile)) {
    (reloo <- readRDS(reloofile))
} else {
    Sys.time()
    (reloo <- reloo(A4, loo, chains=CHAINS, cores=CORES) )
    Sys.time()
    saveRDS(reloo, reloofile)
}
```
Note that the reloo results are cached, and included in the docker image by default.
If you want to reproduce the reloo manually, set a separate cache dir (and mount it into the docker image), and enable the RELOO environment variable.

```{r plot_reloo-a4, eval=params$reloo}
plot(reloo)
```

Running reloo, which does exact LOO-CV (leaving out one problematic observation at a time) reveals that all Pareto-k estimates (the highest is slightly above 0.5) are below 0.7, so we should be able to use the model inferences.

```{r}
pp_check(A4, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(A4, type = "bars_grouped", group="id", ndraws = 500, size = 1/2, fatten = 3/2)
```

Posterior checks look OK.

```{r}
summary(A4)
```

```{r}
plot_M4_latent_distribution(A4, "accessibility")
```

Quite similar distribution for the two sites.

```{r}
plot_M4_posterior_mean(A4, "accessibility-varying", accessible_mean, limits=c(1,7))
```


# Intent to use

## Intent, intercept-only
```{r}
ai_intent <- get_ai_intend_use(ai)
```


```{r}
d <- ai_intent |> select(intent)
formula <- intent | thres(6) ~ 1
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r i1}
I1 <- brm(
  data = d,
  file = modelfile("I1"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- I1
```

No divergent transitions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

The chains seem to mix well, and the `disc` is flat.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is about 0.4.

```{r loo-i1}
( loo <- loo(I1) )
plot(loo)
```

All Pareto-k values are well below 0.7, the highest one around 0.4.

```{r}
pp_check(I1, type = "bars",
         ndraws = 500, size = 1/2, fatten = 3/2)
```

Good LOO-CV and posterior checks.

```{r}
plot_M1_latent_distribution(I1, "intent")
```

```{r}
pp_check(I1, type = "hist", ndraws = 8, binwidth = 1) +
  scale_x_continuous(breaks = 1:7) +
  ggtitle("Example draws from the posterior")
```

```{r}
plot_M1_posterior_mean(I1, "intent",  intent_mean)
```

Estimate is quite close to SL, even though sample mean is a bit closer towards QL.

```{r}
ai_intent |> summarise(mean(as.numeric(intent)))
```

## Intent, per site

```{r}
d <- ai_intent |> select(intent, site)
formula <- intent | thres(6) ~ 1 + site
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r i2}
I2 <- brm(
  data = d,
  file = modelfile("I2"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- I2
```

No divergent transitions.

```{r}
pp_check(m, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```


```{r}
plot_M2_latent_distribution(I2, "intent")
```

A bit more separated distributions, India have a higher tendency towards XL/QL.

```{r}
plot_M2_posterior_mean(I2, "intent",  intent_mean)
```

Which is also visible in the expected value.

## Intent, per site, regularized per respondent

```{r}
d <- ai_intent |> select(intent, site, id)
formula <- intent | thres(6) ~ 1 + site + (1 | id)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(exponential(5), class = sd))
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r i3}
I3 <- brm(
  data = d,
  file = modelfile("I3"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1
)
m <- I3
```

No divergent transitions.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

Chains mix well.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Good diagnostics, lowest $N_{eff}$ is around 0.3.

```{r}
pp_check(m, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(m, type = "bars_grouped", group="id", ndraws = 500, size = 1/2, fatten = 3/2)
```

Good posterior checks

```{r}
plot_M2_latent_distribution(I3, "intent")
```

Similar separation as the non-grouped version I2.

```{r}
plot_M2_posterior_mean(I3, "intent",  intent_mean)
```

## Intent-to-use, pooled on respondent, varying variance per site
```{r}
d <- ai_intent |> select(intent, site, id)
formula <- bf(intent | thres(6) ~ 1 + site + (1 | id)) +
           lf(disc              ~ 0 + site, cmc = FALSE)
priors <- c(prior(normal(-1.068, 1), class = Intercept, coef = 1),
            prior(normal(-0.566, 1), class = Intercept, coef = 2),
            prior(normal(-0.180, 1), class = Intercept, coef = 3),
            prior(normal( 0.180, 1), class = Intercept, coef = 4),
            prior(normal( 0.566, 1), class = Intercept, coef = 5),
            prior(normal( 1.068, 1), class = Intercept, coef = 6),
            prior(normal(0, 1), class = b),
            prior(normal(0, log(2)/2), class = b, dpar=disc),
            prior(exponential(5), class = sd)
            )
validate_prior(prior = priors,
               data=d,
               family=cumulative(probit),
               formula=formula)
```

```{r i4}
I4 <- brm(
  data = d,
  file = modelfile("I4"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
m <- I4
```

No divergent transtions, or starting point problems when we fixed the `init` value and step size `adapt_delta`.

```{r}
p <- mcmc_trace(m)
pars <- levels(p[["data"]][["parameter"]])
plots <- seq(1, to=length(pars), by=6)
lapply(plots, function(i) {
  start <- i
  end <- start+5
  mcmc_trace(m, pars = na.omit(pars[start:end]))
  })
```

Chains mix well for all parameters, including `b_disc_siteIndia`.

```{r}
mcmc_plot(m, type="rhat")
mcmc_plot(m, type="rhat_hist")
mcmc_plot(m, type="neff")
mcmc_plot(m, type="neff_hist")
```

Diagnostics look good, lowest $N_{eff}$ is above 0.25.

```{r loo-i4}
(loo <- loo(I4) )
plot(loo)
```

No problems with the LOO-CV, all Pareto-k values are below 0.7.

```{r}
pp_check(I4, type = "bars", ndraws = 500, size = 1/2, fatten = 3/2)
```

```{r}
pp_check(I4, type = "bars_grouped", group="id", ndraws = 500, size = 1/2, fatten = 3/2)
```

Good posterior checks



```{r}
plot_M4_latent_distribution(I4, "intent-varying")
```

Quite similar distributions between the sites.

```{r}
plot_M4_posterior_mean(I4, "intent-varying",  intent_mean, limits=c(1,7))
```

## Simulating rave reviews

With this model, we can simulate what would happen if we get a new bunch of very positive reviews on one site.

```{r}
I4_fake <- brm(
  data = rbind(d, data.frame(intent="XL", site="India", id=rep(c("100", "101", "102", "103", "104", "105", "106", "107", "108", "109"), 2))),
  file = modelfile("I4-fake-india"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
plot_M4_latent_distribution(I4_fake, "intent-fake-india")

plot_M4_posterior_mean(I4_fake, "intent-fake-india",  intent_mean, limits=c(1,7))
```

Because we have few Europe respondents, when we add 10 Indian respondents, the credible interval for Europe gets wider, and almost reaches to Slightly UnLikely (whereas the Indian cohort gets closer to Extremely Likely).
Europe serves as the reference category, its curve will always be $\mathcal{N}(0,1)$. But the Indian distribution changes, via the dispersion factor.

```{r}
I4_fake <- brm(
  data = rbind(d, data.frame(intent="XL", site="Europe", id=rep(c("100", "101", "102", "103", "104", "105", "106", "107", "108", "109"), 2))),
  file = modelfile("I4-fake-Europe"),
  family = cumulative(probit),
  formula=formula,
  prior = priors,
  drop_unused_levels = F,
  backend="cmdstanr",
  warmup = 1000,
  iter  = ITERATIONS,
  chains = CHAINS,
  cores = CORES,
  seed = 1,
  control = list(adapt_delta=0.99),
  init = 0
)
plot_M4_latent_distribution(I4_fake, "intent-fake-europe")

plot_M4_posterior_mean(I4_fake, "intent-fake-europe",  intent_mean, limits=c(1,7))
```

# Model comparison
```{r}
loo(M1, M2, M3, M4)
```


```{r}
loo(E1, E2, E3, E4)
```


```{r}
loo(A1, A2, A3, A4)
```

```{r}
loo(I1, I2, I3, I4)
```

For the accessibility and intent to use, there is very little difference between the models.
The `elpd_diff` is barely one standard deviation away between the highest and the lowest ranked models.

```{r}
expected_value_M4(M4)
expected_value_M4(E4)
expected_value_M4(A4)
expected_value_M4(I4)
```
